Random Forests

### 2020-07-02 pkg/partykit/inst/ULGcourse-2020/slides_forest.Rnw
further revisions on section 'implementations'
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/partykit@3033 255f4249-2263-4417-b1ea-8498779e27a7

Former-commit-id: d53b48c8616ccb0bbd4e12b91a21b8d609064ccc

\begin{frame}
\frametitle{Variable importance}
{\bf Idea:} Asses the influence of a covariate by evaluating the change in performance of the model (mean decrease in accuracy) after dropping or permutating the selected covariate.

@@

Depending on the fitted model the corresponding objective function can be employed as \emph{risk function} to measure accuracy by calculating importance scores (e.g., log-likelihood, nr. of misclassifications, \ldots).
%}
\end{frame}

@@

<<call_ranger_predict, eval=TRUE, echo=TRUE, results=verbatim>>=
newdata <- data.frame(education = 13, experience = 17,
                      age = 36, ethnicity = "cauc",
                      region = "other", gender = "male",
                      occupation = "worker", sector = "other",
                      union = "no", married = "yes")
pred_rf <- predict(rf, data = newdata)
pred_rf$predictions     
exp(pred_rf$predictions)     
@

@@

\only<1>{
Optional forest-specific control arguments:
\begin{itemize}
\item \code{ytrafo}: transformation function applied to the response,
\item \code{ntree}: number of trees, 
\item \code{mtry}: number of randomly sampled split variable candidates, 
\item \code{perturb}: type and size of subsetting, 
\item \ldots 
\end{itemize}
%\vspace{0.9cm}
}
\only<2->{
Optional tree-specific arguments handed over via \fct{ctree\_control}:
\begin{itemize}
\item \code{minsplit}: minimum number of observations to perform a split, 
\item \code{minbucket}: minimum number of observations in a node after splitting, 
\item \code{maxdepth}: maximum depth of the tree, 
\item \code{alpha}: significance level for split variable selection,
\item \ldots 
\end{itemize}
%\vspace{0.9cm}
}

% \only<3>{
% <<call_cforest_control, eval = FALSE, echo=TRUE>>=
% cf <- cforest(log(wage) ~ education + experience + age + ethnicity + gender + union,
%              data = CPS1985, control = ctree_control(minsplit = 20))
% @
% }

@@

<<call_cforest_predict, eval = TRUE, results=verbatim>>=
newdata <- data.frame(education = 13, experience = 17,
                      age = 36, ethnicity = "cauc",
                      region = "other", gender = "male",
                      occupation = "worker", sector = "other",
                      union = "no", married = "yes")
pred_cf <- predict(cf, newdata = newdata)
pred_cf
exp(pred_cf)
@

>>>>

\begin{frame}
\frametitle{Variable importance}
{\bf Idea:} Asses the influence of a covariate on the model by evaluating the change in performance of the model (mean decrease in accuracy) after permutating the selected covariate.

@@

Depending on the fitted model the corresponding objective function can be employed as \emph{risk function} to measure accuracy by calculating importance scores (e.g., log-likelihood, number of misclassifications, \ldots).
%}
\end{frame}

@@

<<call_ranger_predict, eval=TRUE, echo=TRUE, results=verbatim>>=
newdata <- data.frame(education = 13, experience = 17,
                      age = 36, ethnicity = "cauc",
                      region = "other", gender = "male",
                      occupation = "worker", sector = "other",
                      union = "no", married = "yes")
pred_rf <- predict(rf, data = newdata)
pred_rf$predictions     
exp(pred_rf$predictions) 
rf <- ranger(log(wage) ~ education + experience + age + ethnicity + gender + union,
             data = CPS1985, importance = "impurity")
importance(rf)
@

@@

\only<1>{
Optional forest-specific control arguments:
\begin{itemize}
\item \code{ytrafo}: transformation function applied to the response,
\item \code{ntree}: number of trees, 
\item \code{mtry}: number of randomly sampled split variable candidates, 
\item \code{perturb}: type and size of subsetting, 
\item \ldots 
\end{itemize}
%\vspace{0.9cm}
}
\only<2->{
Optional tree-specific arguments handed over via \fct{ctree\_control}:
\begin{itemize}
\item \code{minsplit}: minimum number of observations to perform a split, 
\item \code{minbucket}: minimum number of observations in a node after splitting, 
\item \code{maxdepth}: maximum depth of the tree, 
\item \code{alpha}: significance level for split variable selection,
\item \ldots 
\end{itemize}
}
% \only<3>{
% 
% <<call_cforest_control, eval = FALSE, echo=TRUE>>=
% cf <- cforest(log(wage) ~ education + experience + age + ethnicity + gender + union,
%             data = CPS1985, control = ctree_control(minsplit = 20))
% @
% }

@@

Optional tree-specific arguments handed over via \fct{ctree\_control}:
\begin{itemize}
\item \code{minsplit}: minimum number of observations to perform a split, 
\item \code{minbucket}: minimum number of observations in a node after splitting, 
\item \code{maxdepth}: maximum depth of the tree, 
\item \code{alpha}: significance level for split variable selection,
\item \ldots 
\end{itemize}

<<call_cforest_control, eval = FALSE, echo=TRUE>>=
cf <- cforest(log(wage) ~ education + experience + age + ethnicity + gender + union,
            data = CPS1985, control = ctree_control(minsplit = 20))
@

\end{frame}
\begin{frame}[fragile]
\frametitle{R software for forest models}

<<call_cforest_predict, eval = TRUE, results=verbatim>>=
newdata <- data.frame(education = 13, experience = 17,
                      age = 36, ethnicity = "cauc",
                      region = "other", gender = "male",
                      occupation = "worker", sector = "other",
                      union = "no", married = "yes")
pred_cf <- predict(cf, newdata = newdata)
pred_cf
exp(pred_cf)
varimp(cf)
@
